---
title: "2024 Infodengue Sprint"
output: html_notebook
---

This script shows how to fit a simple forecast model to the dengue data and upload it to 
the Mosqlimate platform. Your job is to do better than this model!  

## Getting the system ready

To latter upload your projections to the platform, you will use the mosqlient, a python
interface. To run the commands, the reticulate library is required.

```{r}
# un/comment if you need to install these libraries
#install.packages('pak') # useful for installing reticulate
library(pak)
#pak::pak("rstudio/reticulate")
```

Now it is time to install the mosqlient

```{r}
# un/comment if you need to install them
library(reticulate)
#py_install("mosqlient>=1.2.2", envname='r-reticulate')
mosq <- import("mosqlient")
```


## Reading the data

To access the data go to info.dengue.mat.br/minio, with your credentials and 
download the datasets. They are gzip compressed csv files. Here, we will use only
the disease data.  

The dengue data is provided by week and by municipality. 

```{r, results='hide'}
library(data.table)
data <- fread("~/Data/sprint/dengue.csv.gz")
```

```{r}
head(data)
```


Let's for the sake of this example fit the dengue series for the state of Rio de Janeiro (UF="RJ"). For this we need to aggregate the cases from all 92 municipalities of the state.

The first challenge is to forecast 22/23 season using previous data. 

```{r}
dRJ.train <- data %>%
  filter(train_1) %>%
  filter(uf == "RJ") %>% 
  group_by(date) %>% 
  summarize(casos=sum(casos)) %>% 
  arrange(date)
```

```{r}
dRJ.target <- data %>%
  filter(target_1) %>%
  filter(uf == "RJ") %>% 
  group_by(date) %>% 
  summarize(casos=sum(casos)) %>% 
  arrange(date)
```

## Creating a simple forecast model for the challenge 1

For this demonstration we will train an ARIMA model. 

```{r, results='hide'}
library(forecast)
library(tidyverse)
library(ggplot2)
```

First, creating time series objects

```{r}
ts.train <- ts(dRJ.train$casos, start = c(2010, 01), frequency = 52)
ts.target <- ts(dRJ.target$casos, start = c(2022, 41), frequency = 52)
```

Plotting the challenge

```{r}
autoplot(ts.train, colour="blue", xlim = as.Date(c("2010-01-01","2023-10-01")))+
  autolayer(ts.target)  

```

Training the (very bad) arima model

```{r}
fit.train = auto.arima(ts.train)
summary(fit.train)
```
```{r}
autoplot(forecast(fit.train,h=52))+
autolayer(ts.train) + 
  autolayer(ts.target)
```

## Preparing the output for uploading

First, register your model in the platform:

```{r}
library(reticulate)
mosq <- import("mosqlient")
```

You need your API key, to get it, see the instructions in the github readme.

```{r}
library(httr)
url <- "https://api.mosqlimate.org/api/registry/models/"
key = c("your key")
names(key) <- 'X-UID-Key'
model <- list(
    name = "autoarima",
    description = "demo",
    repository = "https://github.com/Mosqlimate-project/sprint-template/",
    implementation_language = "R",
    disease = "dengue",  
    spatial = FALSE,
    temporal = TRUE,
    categorical = FALSE,
    ADM_level = 1,
    time_resolution = "week"
  )
response <- POST(url, body = model, add_headers(.headers=key),  encode = "json")
response
```


Now, to register your prediction, you need: 
1. the id number of your model, see instructions in the github readme
2. dataframe with the following columns: 

date(YYYY-mm-dd): same used in the dataset
preds (num): point estimates
lower (num): 95% lower interval
upper (num): 95% upper interval
adm_1 (two letters string): UF code (example: "RJ")
adm_2: NA  
adm_3: NA


```{r}
uf = "RJ"  # change that accordingly to your data
x <- data.frame(forecast(fit.train,h=52))[,c("Point.Forecast","Lo.95","Hi.95")] 
x <- x %>%
  mutate(dates = as.Date(date_decimal(as.numeric(row.names(x)))),
         adm_1 = uf, 
         adm_2 = NA,
         adm_0 = NA) %>%
  select(dates, 
         preds = Point.Forecast,
         lower = Lo.95,
         upper = Hi.95,
         adm_1,
         adm_2,
         adm_0) %>%
  remove_rownames()

head(x)
```

Converting to JSON

```{r}
library(jsonlite)

y <- toJSON(x, na='string')
y
```


Now, to upload the prediction we will use the `mosqlient` library to handle the POST, instead of the way we did for the model's registration. This will make things substantially easier. For this step you need your *model's id number* found in your profile page in the mosqlimate API website, and collect the commit number of the version of the code at the time of the prediction's generation. 

```{r}
mosq$upload_prediction(
    model_id=15,
    description="arima forecast for RJ",
    commit="505d1b01a5d9ae59a62e635224244c93eef1a466",
    predict_date="2024-06-25",
    prediction=y,
    api_key="your key"
)
```

If you get a 201 status on this submission, it means you are done! Congratulations!
